CYO Project Report - Brazilian Ecommerce

### 2021-07-03 draftReport.Rmd
basic analysis using graphs

```{r Install or load packages}
# TODO use conditional loading or installation of library
library(dplyr)
library(lubridate)
library(tidyverse)
install.packages("geosphere")
library(geosphere)
library(caret)
```

@@

# orders data with item related seller info
# order_item_id --->sequential number identifying number of items included in the same order.
# so if there are three items they will be nuber 1 , 2, 3 for the given order
# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(order_id=as.factor(order_id),product_id=as.factor(product_id),seller_id=as.factor(seller_id), shipping_limit_date=as_datetime(shipping_limit_date))

@@

@@

# identify if there was delay in delivery to carrier
raw_data <- raw_data %>% select(shipping_limit_date,handed_over_to_carrier_date) %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))

```

>>>>

```{r Install or load packages}
# TODO use conditional loading or installation of library
if (!require('dplyr')) install.packages('dplyr'); library('dplyr')
if (!require('lubridate')) install.packages('lubridate'); library('lubridate')
if (!require('tidyverse')) install.packages('tidyverse'); library('tidyverse')
if (!require('geosphere')) install.packages('geosphere'); library('geosphere')
if (!require('caret')) install.packages('caret'); library('caret')
if (!require('plotly')) install.packages('plotly'); library('plotly')
```

@@

# orders data with item related seller info
# order_item_id --->sequential number identifying number of items included in the same order.
# so if there are three items they will be nuber 1 , 2, 3 for the given order
# shipping_limit_date ---> Shows the seller shipping limit date for handling the order over to the logistic partner.
# freight_value --> item freight value item (if an order has more than one item the freight value is splitted between items)
order_items_dataset <- read.csv(file.path('data','olist_order_items_dataset.csv'),header=TRUE,stringsAsFactors = FALSE) %>% mutate(shipping_limit_date=as_datetime(shipping_limit_date))

@@

```{r RMSE}
RMSE <- function(true_ratings, predicted_ratings) {
  sqrt(mean((true_ratings - predicted_ratings)^2))
}
```

@@

# identify if there was delay in delivery to carrier
raw_data <- raw_data %>% mutate(delayed_to_carrier=if_else((as.numeric(handed_over_to_carrier_date)-as.numeric(shipping_limit_date))>0,TRUE,FALSE))

wrangled_data <- raw_data %>% select(-shipping_limit_date,-handed_over_to_carrier_date,-payment_approved_at,-order_id,-order_estimated_delivery_date,-order_item_id,-delayed_to_carrier) %>% mutate(order_purchase_timestamp=as.numeric(order_purchase_timestamp),order_delivered_customer_date=as.numeric(order_delivered_customer_date),product_id=as.factor(product_id),seller_id=as.factor(seller_id),seller_zip_code_prefix=as.factor(seller_zip_code_prefix),customer_unique_id=as.factor(customer_unique_id),customer_zip_code_prefix=as.factor(customer_zip_code_prefix),product_category_name=as.factor(product_category_name))

# we will predict number of hours to deliver based on various predictors. This seems more appropriate. Though as a result of this, we won't be able to use night, day time data of purchase which may influence delivery time (say to carrier it is informed only next day)
wrangled_data <- wrangled_data %>% mutate(number_of_hours=(order_delivered_customer_date-order_purchase_timestamp)/3600) %>% select(-order_delivered_customer_date,-order_purchase_timestamp) 
```

```{r Explore product_category_name relation with number_of_hours}
wrangled_data %>% group_by(product_category_name) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_category_name=reorder(product_category_name,mean)) %>% ggplot(aes(x=product_category_name,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
We can see that number_of_hours of delivery varies as number product category changes

```{r Explore price relation with number_of_hours}
# first let see how are the items associated with orders priced. Most of the order items a
p1 <- wrangled_data %>% mutate(price_group=round(price/100)+1) %>% ggplot(aes(price_group)) + geom_histogram(binwidth=1) + xlab('Each price_group is of 100 units each')
ggplotly(p1)

# let's plot errorbar for number of hours vs price_group.
# we can see while for price groups number of hours for delivery vary but the relationship is not clear enough. Higher priced items can also be delivered early
wrangled_data %>% mutate(price_group=round(price/50)+1) %>% mutate(price_group=as.factor(price_group)) %>% group_by(price_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=price_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(wrangled_data$price,wrangled_data$number_of_hours)
```
Price Group for a product does not give any clear indication. It does not seem to be impacting delivery time of product.

```{r Explore distance relationship between two points with number_of_hours}
# let's plot errorbar for number of hours vs distance group
# we can see a pattern that as distance increases delivery time increases. 
wrangled_data %>% mutate(distance_group=round(distance_between_two_points/50)+1) %>% mutate(distance_group=as.factor(distance_group)) %>% group_by(distance_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=distance_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(wrangled_data$distance_between_two_points,wrangled_data$number_of_hours)
```

Distance Group can be seen to influence delivery time.

```{r Explore freight value group with number_of_hours}
# let's plot errorbar for number of hours vs freight value group
wrangled_data %>% mutate(freight_value_group=round(freight_value/5)+1) %>% mutate(freight_value_group=as.factor(freight_value_group)) %>% group_by(freight_value_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=freight_value_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))

cor(wrangled_data$freight_value,wrangled_data$number_of_hours)
```

Above graphs indicates that for freight value group where each group represents 5 units, as freight value increases, number of hours for delivery increases.

```{r Explore seller_zip_code_prefix with number_of_hours}
# TODO anyway to add regularization or remove the low impact seller_zip_code_prefix from the model
# let's plot errorbar for number of hours vs seller_zip_code_prefix
sellers_zip_of_interest <- wrangled_data %>% dplyr::count(seller_zip_code_prefix) %>% filter(n>10) %>% pull(seller_zip_code_prefix)

wrangled_data %>% filter(seller_zip_code_prefix %in% sellers_zip_of_interest) %>% group_by(seller_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(seller_zip_code_prefix=reorder(seller_zip_code_prefix,mean)) %>%  ggplot(aes(x=seller_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt() 
```
Indicates role between seller_zip_code and number of hours for delivery

```{r Explore customer_zip_code_prefix with number_of_hours}
# TODO anyway to add regularization or remove the low impact customer_zip_code_prefix from the model
# let's plot errorbar for number of hours vs customer_zip_code_prefix
customers_zip_of_interest <- wrangled_data %>% dplyr::count(customer_zip_code_prefix) %>% filter(n>=5) %>% pull(customer_zip_code_prefix)

wrangled_data %>% filter(customer_zip_code_prefix %in% customers_zip_of_interest) %>% group_by(customer_zip_code_prefix) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(customer_zip_code_prefix=reorder(customer_zip_code_prefix,mean)) %>%  ggplot(aes(x=customer_zip_code_prefix,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90)) + scale_y_sqrt()
```
customer_zip_code_prefix can  be seen to be  influence number of hours for delivery

```{r Explore product_id with number_of_hours}
# TODO anyway to add regularization or remove the low impact products from the model
# let's plot errorbar for number of hours vs product_id
products_of_interest <- wrangled_data %>% dplyr::count(product_id)  %>% filter(n>=10) %>% pull(product_id)

wrangled_data %>% filter(product_id %in% products_of_interest) %>% group_by(product_id) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% mutate(product_id=reorder(product_id,mean)) %>%  ggplot(aes(x=product_id,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
It influences the number of hours

```{r Explore volume of product group with number_of_hours}
# let's plot errorbar for number of hours vs volume of product group
wrangled_data %>% mutate(volume_group=round((product_length_cm*product_width_cm*product_height_cm)/1000)+1) %>% mutate(volume_group=as.factor(volume_group)) %>% group_by(volume_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=volume_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
It does not influencee much the number of hours

```{r Explore weight of product with number_of_hours}
# let's plot errorbar for number of hours vs weight of product
wrangled_data %>% mutate(weight_group=round((product_weight_g)/1000)+1) %>% mutate(weight_group=as.factor(weight_group)) %>% mutate(weight_group=as.factor(weight_group)) %>% group_by(weight_group) %>% summarise(mean=mean(number_of_hours),se=sd(number_of_hours)/sqrt(n())) %>% ggplot(aes(x=weight_group,y=mean,ymax=mean+2*se,ymin=mean-2*se)) + geom_point()+geom_errorbar() + theme(axis.text.x = element_text(angle=90))
```
Some clues but not strongly correlated.

Split data into training and validation set, where validation set will be used later
```{r Split data into training and validation set}
set.seed(1,sample.kind = 'Rounding')
test_indices <- createDataPartition(wrangled_data$number_of_hours,times=1,p=0.2,list=FALSE)

training_dataset <- wrangled_data[-test_indices,]
validation_dataset <- wrangled_data[test_indices,]
#TODO ensure validation_dataset has products only for concerned product categories which are in training_dataset 
```

We use training_dataset and split into into train and test set. So that training_dataset is sufficient to fine tune the model
```{r Split data for training}
# Let's split training_dataset into two parts
set.seed(1,sample.kind = 'Rounding')
test_indices <- createDataPartition(training_dataset$number_of_hours,times=1,p=0.2,list=FALSE)

train_subset <- training_dataset[-test_indices,]
test_subset <- training_dataset[test_indices,]
#TODO ensure test_subset has products only for concerned product categories which are in train_subset 
```

```{r Linear model}
# trying to reduce coefficients.
# in dimensions we can use volume to decrase predictors

########### only price predictor#############
## only price complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 220.41

## only price,distance_between_two_points complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 206.98

## only price,distance_between_two_points,product_weight_g complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 206.24

## only price,distance_between_two_points,product_weight_g,product_length_cm complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix)  %>% select(price,distance_between_two_points,product_weight_g,product_length_cm))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 206.22

## only price,distance_between_two_points,product_weight_g,product_length_cm,freight_value complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix)  %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 205.44

## only price,distance_between_two_points,product_weight_g,product_length_cm,product_height_cm,freight_value complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,product_height_cm,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,product_height_cm))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 205.43

## only price,distance_between_two_points,product_weight_g,product_length_cm,product_height_cm,freight_value,product_width_cm complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix)  %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,product_height_cm,product_width_cm,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix)  %>% select(price,distance_between_two_points,product_weight_g,product_length_cm,freight_value,product_height_cm,product_width_cm))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 205.41

## only product_category_name complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(product_category_name,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(product_category_name))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 218.89

# if we combine this with other predictors
## only price,distance_between_two_points,product_category_name complete data
lm_fit <- train(number_of_hours~., train_subset %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_category_name,number_of_hours), method='lm')

predictions <- predict(lm_fit,test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% select(price,distance_between_two_points,product_category_name))

RMSE(predictions,(test_subset  %>% select(-customer_unique_id,-seller_id,-product_id,-customer_zip_code_prefix,-seller_zip_code_prefix) %>% pull(number_of_hours) ))
# 204.87
```

To improve model performance consider--> changing character categorical variables to numerical categorical variables like

  product_category_name=as.factor(as.numeric(product_category_name))

use levels from original data for completeness

look for variables which have limited to no results and we can combine them as one categorical variable
Refer: https://www.analyticsvidhya.com/blog/2015/11/easy-methods-deal-categorical-variables-predictive-modeling/
Also read : feature hashing

Existing model algorithms should handle regularization. So we should not be required to handle ourselves.

We can use conditional model i.e. for some use different model and for some use different model just like ensemble.
So basically we could say for these pin codes, delivery time= say mean of delivery time and for others use standard algorthm like random forest etc.
